---
title: "EDA"
editor: visual
date: "`r Sys.Date()`"
format:
    html:
        toc: true
        embed-resources: true
        theme: default
        code-copy: true
        code-line-numbers: true
        number-sections: true
        highlight-style: github
---

# Synopsis

This dataset consists of medical records for patients at a University hospital outpatient clinic specializing in collagen diseases. The data is divided into three tables that can be connected using patient ID numbers. The first table contains basic patient information and diagnosis details, the second table provides information on examination dates, lab results, and thrombosis occurrences, and the third table includes a comprehensive list of laboratory test results. The main goals of this study are to identify patterns for detecting and predicting thrombosis, find temporal patterns related to thrombosis and collagen diseases, and classify collagen diseases accurately.

# Initial questions

1.Can we identify specific laboratory tests or combinations of tests that are strongly associated with thrombosis, and can these tests be used to predict the likelihood of thrombosis in patients?

2.Are there any temporal patterns in laboratory test results or symptom onset that are specific to certain collagen diseases, and can these patterns be used to improve the accuracy of disease classification?

3.How does the age and sex of patients with collagen diseases impact their laboratory test results, and are there any differences in the test results or disease severity between male and female patients in different age groups?

How do different demographic factors, such as age and sex, affect the prevalence of thrombosis in patients with collagen diseases? Are there any specific laboratory test results (e.g., ANA, aCL IgG, or aCL IgM) that show a strong correlation with the occurrence of thrombosis in these patients? Do specific collagen disease subtypes (e.g., C3 C4) exhibit unique temporal patterns in lab results that can help differentiate them from other subtypes?

# Data Ingestion, cleaning, and munging

```{r}
library(tidyverse)
library(lubridate)
library(snakecase)
library(dplyr)

# Load the three datasets
tsumoto_a <- read.csv('TSUMOTO_A.csv')
tsumoto_b <- read.csv('TSUMOTO_B.csv')
tsumoto_c <- read.csv('TSUMOTO_C.csv', fileEncoding = "latin1")

#Clean the column names by converting them to snake_case format
colnames(tsumoto_a) <- to_snake_case(colnames(tsumoto_a))
colnames(tsumoto_b) <- to_snake_case(colnames(tsumoto_b))
colnames(tsumoto_c) <- to_snake_case(colnames(tsumoto_c))


#Join the three tables using the ID column

combined_data <- tsumoto_a %>%
  full_join(tsumoto_b, by = "id") %>%
  full_join(tsumoto_c, by = "id")



tidy_data <- combined_data %>%
  distinct() %>%
  drop_na(id)%>%
  drop_na(examination_date) # Optionally, remove rows with missing IDs



#subset dataset for required columns of analysis
tidy_data <- subset(tidy_data , select=c("id", "sex", "birthday", "diagnosis.x", "examination_date", "a_cl_ig_g", "a_cl_ig_m", "ana", "ana_pattern", "diagnosis.y", "thrombosis",'date'))


# Replace blank values
tidy_data <- tidy_data %>%
  mutate_all(~replace(., . == "", NA))

#keep the most recent row for each ID
tidy_data <- tidy_data %>%
  arrange(desc(date)) %>%
  distinct(id, .keep_all = TRUE)

tidy_data <- select(tidy_data, -date)

# remove rows where NA still exists
tidy_data <- na.omit(tidy_data)


write.csv(tidy_data, file = "tidy_data.csv", row.names = FALSE)
```

# Exploratory Visual Analysis

```{r}
# Load the tidy dataset
tidy_data <- read.csv("tidy_data.csv", header = TRUE)

# Print the structure of the dataset
str(tidy_data)

# Print summary statistics of numerical variables
summary(tidy_data$value)

head(tidy_data)
```

Examine relationships among variables using scatterplots or correlation matrices

```{r}

# Verify the distribution of thrombosis across different diagnoses
# Convert diagnosis.y column into a format that can be easily summarized
tidy_data$diagnosis_clean <- sapply(strsplit(as.character(tidy_data$diagnosis.y), ", "), function(x) x[1])

# Summarize thrombosis by the cleaned diagnosis
summary_table <- tidy_data %>%
  group_by(diagnosis_clean, thrombosis) %>%
  summarize(count = n()) %>%
  ungroup()

print(summary_table)

# Examine the correlation between lab test results (a_cl_ig_g, a_cl_ig_m, ana) and thrombosis occurrence

# Plot the relationship between a_cl_ig_g, a_cl_ig_m, and thrombosis
library(ggplot2)
ggplot(tidy_data, aes(x=a_cl_ig_g, y=a_cl_ig_m, color=factor(thrombosis))) +
  geom_point() +
  labs(title="Relationship between aCL IgG, aCL IgM, and Thrombosis")

# Compute the correlation coefficients between lab test results and thrombosis
cor_matrix <- cor(tidy_data[, c("a_cl_ig_g", "a_cl_ig_m", "ana", "thrombosis")], use="complete.obs")
print(cor_matrix)

```

phase 2

To get a better sense of the distribution of each variable, we can create a histogram or density plot. For example, to create a histogram of the sex variable

```{r}
ggplot(tidy_data, aes(x = sex, fill = thrombosis)) +
  geom_bar() +
  labs(title = "Prevalence of Thrombosis by Sex",
       x = "Sex",
       y = "Count",
       fill = "Thrombosis")
```

This chart shows that there are more female patients than male patients in the dataset, and that the prevalence of thrombosis is higher in female patients than in male patients.

```{r}
# Explore the correlation between lab test results and thrombosis occurrence
ggplot(tidy_data, aes(x=ana, y=a_cl_ig_g, color=thrombosis)) +
  geom_point() +
  labs(title="Relationship between ANA, aCL IgG, and Thrombosis")

# Investigate temporal patterns in lab results for different collagen disease subtypes:
```

In the second phase of the analysis, we investigated three specific research questions using data visualizations. We plotted the prevalence of thrombosis by age and sex, the correlation between specific lab results and the occurrence of thrombosis, and the temporal patterns of lab results for each collagen disease subtype. Our findings can be used to gain insights into the relationships among the variables and to inform future research.

# Generate publication quality plots

```{r}

```

# Technical summary

```{r}

```

# Non-technical report

```{r}

```
